Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-22 cav2017/cav2017.tex
text

We investigate two different approaches of integrating solution generalisation 
with CEGIS:
\begin{itemize}
\item[1.] The first approach (described in Sec.~\ref{sec:CEGIS-precision-incrementation})
  uses a multi-staged verification phase and performs solution generalisation based on
the plant precision, the number of loop unrollings, 
and the sampling rate. Essentially, it finds a controller for an individual initial state
and input with a bounded time horizon and generalises to all states,
inputs, and time horizons. %% (with a final verification loop that verifies
%% the continuous system behaviour against the previously verified
%% discrete system bi-simulation)
\item[2.] The second approach (described in Sec~\ref{sec:CEGIS-abstraction-refinement})
  %% uses abstract acceleration to compute the set of reachable states,
  %% which enables it to perform abstraction refinement inside the CEGIS loop.
  finds a controller for a continuous initial set of states and input set, 
  over an abstraction of the continuous dynamics that conforms to witness proofs at specific times; 
  it further generalises the solution over all times by refining the number of witnesses. 
\end{itemize}

@@

The verification of a hybrid continuous/discrete system over an unbounded time, requires the
non-deterministic simulation of an infinite set of states, which is undecidable. Our initial approach creates an
abstraction of this system by establishing bounds on the error between such a system and a discrete-time/continuous-space model.
This model is still however, quite complex, hard to verify, and many of the states involved are inconsequential to
the safety specification.
Our second model, rather than look at a bisimulation in discrete time, explores a much higher abstraction,
evaluating the unbounded time-continuous space reach tube as a set of inequalities that represent only the safety specifcation.
We begin with a minimum set of linear inequalities describing the constraints caused by the input and the initial state during a
single iteration, in addition to a stability specification defined by Jury's criteria, and try to synthesise a controller that meets these specs.
A second abstraction that encompasses the accelerated continuous reach-tube over an unbounded time is used to verify that the controller meets the full safety specification. If this step fails, we find a counterexample iteration and initial state and create a new constraint to refine our abstraction.
The advantage of this model is that it starts with a very simple description regardless of the complexity of the overall dynamics, and only expands to more complex models when a solution cannot be found.

@@

Let us recall the formulas used for single-input and single-output (SISO) systems 
in sections \ref{sec:reachable} and \ref{sec:observable}.
A controllable system with quantization will have the following dynamics:

%
\begin{align}
\label{eq:observer_LTI_cf}
\vec{x}_{k+1}=&\mat{A}_{t}\vec{x}_k+\mat{B}_{tp} \vec{r}_k+\mat{B}_{tn} (\vec{\nu}_1 + \vec{\nu}_2 + \vec{\nu}_3)
\end{align}
\begin{align*}
\mat{A}_{t}=&\mat{T}\mat{A}_d\mat{T}^{-1}&
\mat{B}_{tp}=&\mat{T}\mat{B}_d&
\mat{B}_{tn}=&\mat{T}\mat{B}_d&
\mat{T}=&\mat{R}_{cf}\mat{R}^{-1}&
\end{align*}

@@

\begin{figure}
\centering
{\scriptsize
\hspace*{-1cm}
\begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center},row sep=.8cm, column sep=.2cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \& ;\\
   \&\node[fill=gray!20,align=center] (pre) {{\sc 1. pre-}\\{\sc processing}};
   \& \node[fill=gray!20,align=center] (abstract) {\sc 4. abstraction};
       % \node[draw=none] (FIK) at ([xshift=2.5cm,yshift=.3cm]abstract)  {\sc $\Psi (\mat{K})$};
   \& \coordinate (aux); \\ % \node[fill=yellow!20,align=center] (counter) {\sc counterexample};\\
   \& 
   \& \node[fill=gray!20,align=center, minimum width=3cm] (synth) {\sc 2. synthesize};
       %\node[draw=none] (KL) at ([xshift=2.5cm,yshift=-.25cm]synth)  {\sc $\mat{K},\mat{L}$};
       %\node[draw=none] (NKL) at ([xshift=2cm,yshift=1cm]synth)  {\sc $P_a$};
   \& \node[fill=gray!20,align=center, minimum width=4cm] (verify) {\sc 3. verify ($\Psi$)};
       %\node[draw=none] (VUNS) at ([xshift=.6cm,yshift=1cm]verify)  {\sc SAT};
       \node[draw=none] (SAT) at ([xshift=2.7cm,yshift=-.3cm]verify)  {\sc UNSAT};
   \& \node[ellipse, fill=gray!20] (done) {{\sc Done}};\\
   \&
   \& complexnode/.pic={
     \coordinate (auxsynth);
     \node[draw,rectangle,align=center] (KSAT) at ([xshift=-.8cm]auxsynth.center) {\sc Prog.\\Search};
     \node[draw=none] (SATK) at ([xshift=-.5cm,yshift=1.5cm]KSAT)  {$\mathcal{J}_K$};
     \node[draw=none] (K) at ([xshift=.4cm,yshift=1.5cm]KSAT)  {$\tilde{\mat{K}}$};
    }
   \& complexnode/.pic={
     \coordinate (AA);
     \node[draw,rectangle,align=center] (AAI) at ([xshift=-1.5cm]AA.center) {\sc AI\\Model};
     \node[draw,rectangle,align=center] (AAV) at ([xshift=1.5cm]AA.center) {\sc AA\\Solver};
     \node[draw=none] (AB) at ([xshift=.4cm,yshift=1.5cm]AAI.center)  {\sc $\mat{A}_t$ $\mat{B}_{tp}$ $\mat{B}_{tn}$};     
     \node[draw=none] (MA) at ([yshift=.2cm]AA.center)  {$\mathcal{A}_t$};     
     \node[draw=none] (MB) at ([yshift=-.3cm]AA.center)  {$\mathcal{B}_{tp},\mathcal{B}_{tn}$}; 
     \node[draw=none] (AB) at ([xshift=0.3cm,yshift=1.5cm]AAV.center)  {\sl $\hat{X}^\#$};  
    }\\
   \&
   \& complexnode/.pic={
     \coordinate (UNSAT);
     \coordinate (KUNSAT)  at ([xshift=-.8cm]UNSAT.center);
     \node[draw=none] (UNSATB) at ([xshift=1.6cm,yshift=.3cm]UNSAT)  {\sc UNSAT};
  }
  \& \& \coordinate (FAIL);\\
  };
  \path
    (pre.east) edge (abstract.west)
    (abstract.south) edge (synth.north)
    %([yshift=2em]synth.east) edge node {Candidate} ([yshift=2em]verif.west)
    %([yshift=-2em]verif.west) edge node {New input} ([yshift=-2em]synth.east)
    %([yshift=2em]synth.east) edge node[xshift=-0.7em] {Candidate} ([yshift=2em]verif.west)
    %([yshift=-2em]verif.west) edge node[xshift=-0.7em] {New input} ([yshift=-2em]synth.east)
    ([yshift=1em]synth.east) edge node {$\mat{K},\mat{L}$} ([yshift=1em]verify.west)
    ([yshift=-1em]verify.west) edge node {$P_a$} ([yshift=-1em]synth.east)
    (aux) edge (abstract.east) 
    (verify.east) edge (done.west);
  \path
    ([xshift=+.6cm]KSAT.north) edge ([xshift=-2.1cm]synth.south) 
    ([xshift=-3.3cm]synth.south) edge ([xshift=-.6cm]KSAT.north)
    ([xshift=-4.9cm]verify.south) edge (AAI.north)
    (AAI.east) edge (AAV.west)
    (AAV.north) edge ([xshift=4.9cm]verify.south)
    %(verify.north) edge (counter.south)
    %(counter.west) edge (abstract.east)
    %(counter.south west) edge (synth.north east)
    (FAIL.north) edge (done.south);
   \path[-] 
     (KSAT.south) edge (KUNSAT.north)
     (KUNSAT.east) edge (FAIL.west)
     (verify.north) edge node {$\Psi (\mat{K})$} (aux);
\end{tikzpicture}
}
\caption{CEGIS with abstraction refinement\label{fig:CEGARIS}.}
\end{figure}

>>>>

We investigate two different approaches of integrating solution generalisation 
with CEGIS:
\begin{itemize}
\item[1.] The first approach (described in Sec.~\ref{sec:CEGIS-precision-incrementation})
  uses a multi-staged verification phase and performs solution generalisation based on
the plant precision, the number of loop unrollings, 
and the sampling rate. Essentially, it finds a controller for an individual initial state
and input with a bounded time horizon and generalises to all states,
inputs, and time horizons. %% (with a final verification loop that verifies
%% the continuous system behaviour against the previously verified
%% discrete system bi-simulation)
\item[2.] The second approach (described in Sec~\ref{sec:CEGIS-abstraction-refinement})
  %% uses abstract acceleration to compute the set of reachable states,
  %% which enables it to perform abstraction refinement inside the CEGIS loop.
  finds a controller for a continuous initial set of states and input set, 
  over an abstraction of the continuous dynamics that conforms to witness proofs at specific times; 
  it further generalises the solution over all times by refining the number of witnesses.  \todo{clarify use of abstract acceleration. add references.}
\end{itemize}

@@

The verification of a hybrid continuous/discrete system over an unbounded time, requires the
non-deterministic simulation of an infinite set of states, which is undecidable. Our initial approach creates an
abstraction of this system by establishing bounds on the error between such a system and a discrete-time/continuous-space model.
This model is still however, quite complex, hard to verify, and many of the states involved are inconsequential to
the safety specification.
Our second model, rather than look at a bisimulation \todo{specific notion that has not been formally define - do not use} in discrete time, explores a much higher abstraction, 
evaluating the unbounded time-continuous space reach tube as a set of inequalities that represent only the safety specification.
We begin with a minimum set of linear inequalities describing the constraints caused by the input and the initial state during a
single iteration, in addition to a stability specification defined by Jury's criteria, and try to synthesise a controller that meets these specs.
A second abstraction that encompasses the accelerated continuous reach-tube over an unbounded time is used to verify that the controller meets the full safety specification. If this step fails, we find a counterexample iteration and initial state and create a new constraint to refine our abstraction.
The advantage of this model is that it starts with a very simple description regardless of the complexity of the overall dynamics, and only expands to more complex models when a solution cannot be found.

@@

Let us recall the formulas used for single-input and single-output (SISO) systems 
in sections \ref{sec:reachable} and \ref{sec:observable}.
A controllable system with quantization will have the following dynamics:
\todo{eliminate vectors}
%
\begin{align}
\label{eq:observer_LTI_cf}
\vec{x}_{k+1}=&\mat{A}_{t}\vec{x}_k+\mat{B}_{tp} \vec{r}_k+\mat{B}_{tn} (\vec{\nu}_1 + \vec{\nu}_2 + \vec{\nu}_3)
\end{align}
\begin{align*}
\mat{A}_{t}=&\mat{T}\mat{A}_d\mat{T}^{-1}&
\mat{B}_{tp}=&\mat{T}\mat{B}_d&
\mat{B}_{tn}=&\mat{T}\mat{B}_d&
\mat{T}=&\mat{R}_{cf}\mat{R}^{-1}&
\end{align*}

@@

\begin{figure}
\centering
{\scriptsize
\hspace*{-1cm}
\begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center},row sep=.8cm, column sep=.2cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \& ;\\
   \&\node[fill=gray!20,align=center] (pre) {{\sc 1. pre-}\\{\sc processing}};
   \& \node[fill=gray!20,align=center] (abstract) {\sc 4. abstraction};
       % \node[draw=none] (FIK) at ([xshift=2.5cm,yshift=.3cm]abstract)  {\sc $\Psi (\mat{K})$};
   \& \coordinate (aux); \\ % \node[fill=yellow!20,align=center] (counter) {\sc counterexample};\\
   \& 
   \& \node[fill=gray!20,align=center, minimum width=3cm] (synth) {\sc 2. synthesize};
       %\node[draw=none] (KL) at ([xshift=2.5cm,yshift=-.25cm]synth)  {\sc $\mat{K},\mat{L}$};
       %\node[draw=none] (NKL) at ([xshift=2cm,yshift=1cm]synth)  {\sc $P_a$};
   \& \node[fill=gray!20,align=center, minimum width=4cm] (verify) {\sc 3. verify ($\Psi$)};
       %\node[draw=none] (VUNS) at ([xshift=.6cm,yshift=1cm]verify)  {\sc SAT};
       \node[draw=none] (SAT) at ([xshift=2.7cm,yshift=-.3cm]verify)  {\sc UNSAT};
   \& \node[ellipse, fill=gray!20] (done) {{\sc Done}};\\
   \&
   \& complexnode/.pic={
     \coordinate (auxsynth);
     \node[draw,rectangle,align=center] (KSAT) at ([xshift=-.8cm]auxsynth.center) {\sc Prog.\\Search};
     \node[draw=none] (SATK) at ([xshift=-.5cm,yshift=1.5cm]KSAT)  {$\mathcal{J}_K$};
     \node[draw=none] (K) at ([xshift=.4cm,yshift=1.5cm]KSAT)  {$\tilde{\mat{K}}$};
    }
   \& complexnode/.pic={
     \coordinate (AA);
     \node[draw,rectangle,align=center] (AAI) at ([xshift=-1.5cm]AA.center) {\sc AI\\Model};
     \node[draw,rectangle,align=center] (AAV) at ([xshift=1.5cm]AA.center) {\sc AA\\Solver};
     \node[draw=none] (AB) at ([xshift=.4cm,yshift=1.5cm]AAI.center)  {\sc $\mat{A}_t$ $\mat{B}_{tp}$ $\mat{B}_{tn}$};     
     \node[draw=none] (MA) at ([yshift=.2cm]AA.center)  {$\mathcal{A}_t$};     
     \node[draw=none] (MB) at ([yshift=-.3cm]AA.center)  {$\mathcal{B}_{tp},\mathcal{B}_{tn}$}; 
     \node[draw=none] (AB) at ([xshift=0.3cm,yshift=1.5cm]AAV.center)  {\sl $\hat{X}^\#$};  
    }\\
   \&
   \& complexnode/.pic={
     \coordinate (UNSAT);
     \coordinate (KUNSAT)  at ([xshift=-.8cm]UNSAT.center);
     \node[draw=none] (UNSATB) at ([xshift=1.6cm,yshift=.3cm]UNSAT)  {\sc UNSAT};
  }
  \& \& \coordinate (FAIL);\\
  };
  \path
    (pre.east) edge (abstract.west)
    (abstract.south) edge (synth.north)
    %([yshift=2em]synth.east) edge node {Candidate} ([yshift=2em]verif.west)
    %([yshift=-2em]verif.west) edge node {New input} ([yshift=-2em]synth.east)
    %([yshift=2em]synth.east) edge node[xshift=-0.7em] {Candidate} ([yshift=2em]verif.west)
    %([yshift=-2em]verif.west) edge node[xshift=-0.7em] {New input} ([yshift=-2em]synth.east)
    ([yshift=1em]synth.east) edge node {$\mat{K},\mat{L}$} ([yshift=1em]verify.west)
    ([yshift=-1em]verify.west) edge node {$P_a$} ([yshift=-1em]synth.east)
    (aux) edge (abstract.east) 
    (verify.east) edge (done.west);
  \path
    ([xshift=+.6cm]KSAT.north) edge ([xshift=-2.1cm]synth.south) 
    ([xshift=-3.3cm]synth.south) edge ([xshift=-.6cm]KSAT.north)
    ([xshift=-4.9cm]verify.south) edge (AAI.north)
    (AAI.east) edge (AAV.west)
    (AAV.north) edge ([xshift=4.9cm]verify.south)
    %(verify.north) edge (counter.south)
    %(counter.west) edge (abstract.east)
    %(counter.south west) edge (synth.north east)
    (FAIL.north) edge (done.south);
   \path[-] 
     (KSAT.south) edge (KUNSAT.north)
     (KUNSAT.east) edge (FAIL.west)
     (verify.north) edge node {$\Psi (\mat{K})$} (aux);
\end{tikzpicture}
}
\caption{CEGIS with abstraction refinement}
\label{fig:CEGARIS} 
\end{figure}
